Syntactic realization with data-driven neural tree grammars

### 2016-10-11 paper/draft.tex
unshrinkage

\vspace*{-0.5cm}
\begin{abstract}

@@

\section{Introduction}
\vspace*{-0.2cm}

@@

\section{Tree Grammars}
\label{sec:tree}
\vspace*{-0.2cm}

@@

\subsection{Grammar Formalism}
\vspace*{-0.1cm}

@@

\subsection{Grammar Induction}
\vspace*{-0.1cm}

@@

\begin{figure*}[tH!]
\centering
\includegraphics[width=\textwidth]{spineembed.pdf}
\caption{Embedding supertags using convolutional neural networks. In (A), a tree is encoded by its features and then embedded.
 In (B), convolutional layers are used to encode the supertag into a vector.}
 \label{fig:spineembedding}
\vspace*{-0.5cm}
\end{figure*}

\section{Neural Representations}
\label{sec:neural}
\vspace*{-0.2cm}

@@

\subsection{Embedding Supertags}
\vspace{-0.1cm}

@@

The simplest way to embed supertags is to treat each structure
as a distinct token that indexes a corresponding learned vector.  This
places no constraints on the learned similarity function, but it also
ignores the hierarchical structure of the elementary trees
themselves.  Previous work on deep learning with graph structures
suggests convolutional neural networks can exploit similarities in structure.  Thus, we developed
analogous techniques to encode supertags based on their underlying
tree structure.  In particular, to embed a supertag, we embed each
node, group the resulting vectors to form a tensor, and then summarize
the tensor into a single vector using a series of convolutional neural
networks.

@@

\vspace*{-0.5cm}
\begin{align}
& T_\Lambda = [G_s(\Lambda_{syntactic~category});~G_r(\Lambda_{role})] \nonumber \\
& G_\Lambda = C^{4,5}(C^{3,1}(C^{1,3}(C^{2,1}(C^{1,2}(T_\Lambda))))) \label{eq:embed}
\end{align}
\vspace*{-0.5cm}

@@

%%%%
\subsection{Recurrent Tree Networks}
\label{subsec:rtn}
\vspace{-0.1cm}

\begin{figure*}[tH!]
\centering
\includegraphics[width=0.8\textwidth]{rtn2.pdf}
\caption{A recurrent tree network. (A) The dependency structure as a tree.  (B) the dependency structure as a sequence.}
 \label{fig:rtn}
\vspace*{-0.5cm}
\end{figure*}

@@

\vspace*{-0.5cm}
\begin{align}
       s_c~&=~RTN(x_t,p_t) \nonumber \\
           &=~g(x_t,S_{tree}[p_t]) \nonumber \\
           &=~g(x_t, s_p) \nonumber \\
S_{tree}[t]&=~s_c \label{eq:rnnsetup}
\end{align}
\vspace*{-0.5cm}

@@

\section{Models}
\label{sec:models}
\vspace*{-0.2cm}

@@

\subsection{Model 1: Fergus-N}
\vspace{-0.1cm}

@@

\vspace*{-0.55cm}
\begin{align}
&min_{\theta} -[\sum_p\sum_{p\to c} log[P_\theta(tag_{p} | lex_{p}, lex_{c}, tag_{c})] + \sum_c log[P_\theta(tag_{c} |lex_{p}, lex_{c})]] \label{eq:fergusn_obj}
\end{align}
\vspace*{-0.45cm}

@@

\vspace*{-0.55cm}
\begin{align}
&\omega_{lex} = FC_1([G_w(lex_p); G_w(lex_c)]) \label{eq:fergusn} \\
&\omega_{node}=concat([G_\Lambda(tag_c);~\omega_{lex}]) \nonumber \\
&\Omega_{treelet} = concat([repeat(\omega_{node}),~G_\Lambda(tagset_p)]) \nonumber \\
&\Omega_{leaf} = concat([repeat(\omega_{lex}),~G_\Lambda(tagset_c)]) \nonumber \\
&P_\theta(tag_{p,i} | lex_{p}, lex_{c}, tag_{c})=
\frac{exp(\Psi_{predict}(\omega_{treelet_i})))}
{\sum_{j \in |tagset_p|} exp(\Psi_{predict}(\omega_{treelet_j})))} \nonumber \\
&P_\theta(tag_{c,i} |lex_{p}, lex_{c}) = 
\frac{exp(\Psi_{predict}(\omega_{leaf_i})))}
{\sum_{j \in |tagset_c|} exp(\Psi_{predict}(\omega_{leaf_j})))} \nonumber 
\end{align}
\vspace*{-0.45cm}

@@

\subsection{Model 2: Fergus-R}
\vspace{-0.1cm}

@@

\vspace*{-0.55cm}
\begin{align}
&min_{\theta} -[\sum_{(p,C)} P_\theta(tag_{p}|RTN,~lex_p,~lex_{C})] \label{eq:rtnobj}
\end{align}
\vspace*{-0.45cm}

@@

\vspace*{-0.55cm}
\begin{align}
&\omega_{C} = \Psi_{attn}(G_w(lex_C)) \label{eq:fergusr} \\
&\omega_{lex} = FC_1(concat(\omega_{C},~G_w(lex_p))) \nonumber \\
&\omega_{node} = RTN(\omega_{lex},~topology) \nonumber \\
&\Omega_{elementary} = FC_2(concat(repeat(\omega_{node}),~G_\Lambda(tagset_p))) \nonumber \\
&P_\theta(tag_{p,i}~|~RTN,~lex_p,~lex_{C}) = 
\frac{exp(\Psi_{predict}(\omega_{elementary_i})))}
{\sum_{j \in |\Omega|} exp(\Psi_{predict}(\omega_{elementary_j))}} \nonumber
\end{align}
\vspace*{-0.45cm}

@@

\subsection{Linearization}
\vspace{-0.1cm}

@@

\section{Experiments}
\label{sec:expt}
\vspace*{-0.2cm}

@@

\subsection{Training}
\vspace{-0.1cm}

@@

\subsection{Performance Metrics}
\vspace{-0.1cm}

@@

\subsection{Results}
\label{sec:results}
\vspace*{-0.1cm}

\begin{table}
\centering
\begin{tabular}{|l|p{3cm}|c|c|c|}
\cline{3-4}
\multicolumn{2}{}{}& \multicolumn{2}{ |c| }{ \textbf{Accuracy}}&\multicolumn{1}{}{} \\ \hline
\textbf{Model} & \textbf{Embedding}  & \textbf{Raw Model} 
& \textbf{After Decoding} & \textbf{Running Time} \\ \hline
Fergus-N &  Structural  &  58.17\% & 57.40\%  & 1.97s \\ \cline{2-5}
         &  Atomic      &  60.69\% & 55.56\% & 1.81s \\ \cline{2-5}
         &  Minimal     &  52.09\% & 54.18\% & 2.02s \\ \cline{2-5}
\hline
Fergus-R &  Structural &  67.62\% & 57.04\% & 0.30s \\ \cline{2-5}
         &  Atomic     &  82.65\% & 62.73\% & 0.36s\\ \cline{2-5}
         &  Minimal    & 10.13\% & 19.66\% & 0.54s \\ \cline{2-5}
\hline
\end{tabular}
\caption{For each supertag and embedding pair, the mean accuracy of
  supertag classification directly output by the model and in the
  consistent global assignment output by A* decoding. Also shown is
  the median running time---which includes model computation and A*
  search. The structural embeddings are computed with convolutional coding, the atomic embeddings as rows in a matrix, and the minimal embeddings as scalars in a vector.}
\label{table:accresults}
\vspace*{-0.5cm}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{|l|p{3cm}|p{2.5cm}|r|}
\cline{3-4}
\multicolumn{2}{}{} & \multicolumn{2}{|c|}{Accuracy}   \\ \hline
\textbf{Model} & \textbf{Embedding}  & \textbf{Top Scoring} & \textbf{Best Performance} \\ \hline
Fergus-N & Structural & 65.80\% & 72.58\% \\ \cline{2-4}
         & Atomic       & 65.52\%  & 71.82\% \\ \cline{2-4}
         & Minimal & 63.79\% & 71.09\% \\ \cline{2-4}
\hline
Fergus-R & Structural & 68.22\% & 74.70\% \\ \cline{2-4}
         & Atomic       &  69.29\% & 75.56\% \\ \cline{2-4}
         & Minimal &  58.23\% & 65.04\% \\ \cline{2-4}
\hline
\end{tabular}
\caption{Shown above as accuracy is the percentage of tokens in the linearized strings that are in correct positions according to an edit distance measure.}
\label{table:linresults}
\vspace*{-0.5cm}
\end{table}

@@

\section{Related Work}
\label{sec:relatedwork}
\vspace*{-0.2cm}

@@

\section{Conclusion}
\label{sec:conclusion}
\vspace*{-0.2cm}

@@

\section*{Acknowledgments}
\vspace*{-0.2cm}

@@

\section{Appendix}
\label{sec:appendix}
\vspace{-0.2cm}

\begin{table}[t!]
\centering
\begin{tabular}{lr}
\begin{tabular}[t]{|p{6cm}|r|}
\hline
\textbf{Model Parameter} & \textbf{Value} \\ \hline
\hline \multicolumn{2}{|c|}{\textbf{Fergus-N Parameters}} \\ \hline
Fully connected layer size & 256 \\ \hline
Batch size & 128 \\ \hline
\hline  \multicolumn{2}{|c|}{\textbf{Fergus-R Parameters}} \\ \hline
Fully connected layer size & 256 \\ \hline 
Hidden state size & 128 \\ \hline
Batch size & 16 \\ \hline
\hline \multicolumn{2}{|c|}{\textbf{Embedding Parameters}} \\ \hline
Convolution filter size & 48 \\ \hline
Syntactic category embedding size & 32 \\ \hline
Node role embedding size & 32 \\ \hline
Word embedding size \newline
\cite{pennington2014glove} & 300 \\ \hline
\end{tabular}
&
\begin{tabular}[t]{|p{6cm}|r|}
\hline
\textbf{Model Parameter} & \textbf{Value} \\ \hline
\hline \multicolumn{2}{|c|}{\textbf{Language Model Parameters}} \\ \hline
Hidden state size & 368 \\ \hline
Batch size & 32 \\ \hline
\hline \multicolumn{2}{|c|}{\textbf{Optimization Parameters}} \\ \hline
Optimization Algorithm & ADAM \\ \hline
Fergus-R and Fergus-N Learning Rate & 1e-4 \\ \hline
Language Model Learning Rate & 0.01 \\ \hline
Fully-Connected Dropout Rate & 0.5 \\ \hline
Recurrent Weight Dropout Rate & 0.2 \\ \hline
$L_2$ Weight Decay & 1e-6 \\ \hline
Max gradient norm & 10.0 \\ \hline
Gradient clip threshold & 5.0 \\ \hline
\end{tabular}
\end{tabular}
\caption{The parameters for the Fergus-R, Fergus-N, and language models.  The
exact specifications in configuration files can be found in the code repository that accompanies this paper.}
\label{tab:params}
\vspace*{-0.5cm}
\end{table}

@@

\subsection{Data}
\vspace*{-0.1cm}

@@

\subsection{Model parameters}
\vspace*{-0.1cm}

@@

>>>>

%\vspace*{-0.5cm}
\begin{abstract}

@@

\section{Introduction}
%\vspace*{-0.2cm}

@@

\section{Tree Grammars}
\label{sec:tree}
%\vspace*{-0.2cm}

@@

\subsection{Grammar Formalism}
%\vspace*{-0.1cm}

@@

\subsection{Grammar Induction}
%\vspace*{-0.1cm}

@@

\begin{figure*}[tH!]
\centering
\includegraphics[width=\textwidth]{spineembed.pdf}
\caption{Embedding supertags using convolutional neural networks. In (A), a tree is encoded by its features and then embedded.
 In (B), convolutional layers are used to encode the supertag into a vector.}
 \label{fig:spineembedding}
%\vspace*{-0.5cm}
\end{figure*}

\section{Neural Representations}
\label{sec:neural}
%\vspace*{-0.2cm}

@@

\subsection{Embedding Supertags}
%\vspace{-0.1cm}

@@

The simplest way to embed supertags is to treat each structure
as a distinct token that indexes a corresponding learned vector.  This
places no constraints on the learned similarity function, but it also
ignores the hierarchical structure of the elementary trees
themselves.  Previous work on deep learning with graph structures
suggests convolutional neural networks can exploit similarities in structure \cite{niepert2016learning,kalchbrenner2014convolutional}.  Thus, we developed
analogous techniques to encode supertags based on their underlying
tree structure.  In particular, to embed a supertag, we embed each
node, group the resulting vectors to form a tensor, and then summarize
the tensor into a single vector using a series of convolutional neural
networks.

@@

%\vspace*{-0.5cm}
\begin{align}
& T_\Lambda = [G_s(\Lambda_{syntactic~category});~G_r(\Lambda_{role})] \nonumber \\
& G_\Lambda = C^{4,5}(C^{3,1}(C^{1,3}(C^{2,1}(C^{1,2}(T_\Lambda))))) \label{eq:embed}
\end{align}
%\vspace*{-0.5cm}

@@

%%%%
\subsection{Recurrent Tree Networks}
\label{subsec:rtn}
%\vspace{-0.1cm}

\begin{figure*}[tH!]
\centering
\includegraphics[width=0.8\textwidth]{rtn2.pdf}
\caption{A recurrent tree network. (A) The dependency structure as a tree.  (B) the dependency structure as a sequence.}
 \label{fig:rtn}
%\vspace*{-0.5cm}
\end{figure*}

@@

%\vspace*{-0.5cm}
\begin{align}
       s_c~&=~RTN(x_t,p_t) \nonumber \\
           &=~g(x_t,S_{tree}[p_t]) \nonumber \\
           &=~g(x_t, s_p) \nonumber \\
S_{tree}[t]&=~s_c \label{eq:rnnsetup}
\end{align}
%\vspace*{-0.5cm}

@@

\section{Models}
\label{sec:models}
%\vspace*{-0.2cm}

@@

\subsection{Model 1: Fergus-N}
%\vspace{-0.1cm}

@@

%\vspace*{-0.55cm}
\begin{align}
&min_{\theta} -[\sum_p\sum_{p\to c} log[P_\theta(tag_{p} | lex_{p}, lex_{c}, tag_{c})] + \sum_c log[P_\theta(tag_{c} |lex_{p}, lex_{c})]] \label{eq:fergusn_obj}
\end{align}
%\vspace*{-0.45cm}

@@

%\vspace*{-0.55cm}
\begin{align}
&\omega_{lex} = FC_1([G_w(lex_p); G_w(lex_c)]) \label{eq:fergusn} \\
&\omega_{node}=concat([G_\Lambda(tag_c);~\omega_{lex}]) \nonumber \\
&\Omega_{treelet} = concat([repeat(\omega_{node}),~G_\Lambda(tagset_p)]) \nonumber \\
&\Omega_{leaf} = concat([repeat(\omega_{lex}),~G_\Lambda(tagset_c)]) \nonumber \\
&P_\theta(tag_{p,i} | lex_{p}, lex_{c}, tag_{c})=
\frac{exp(\Psi_{predict}(\omega_{treelet_i})))}
{\sum_{j \in |tagset_p|} exp(\Psi_{predict}(\omega_{treelet_j})))} \nonumber \\
&P_\theta(tag_{c,i} |lex_{p}, lex_{c}) = 
\frac{exp(\Psi_{predict}(\omega_{leaf_i})))}
{\sum_{j \in |tagset_c|} exp(\Psi_{predict}(\omega_{leaf_j})))} \nonumber 
\end{align}
%\vspace*{-0.45cm}

@@

\subsection{Model 2: Fergus-R}
%\vspace{-0.1cm}

@@

%\vspace*{-0.55cm}
\begin{align}
&min_{\theta} -[\sum_{(p,C)} P_\theta(tag_{p}|RTN,~lex_p,~lex_{C})] \label{eq:rtnobj}
\end{align}
%\vspace*{-0.45cm}

@@

%\vspace*{-0.55cm}
\begin{align}
&\omega_{C} = \Psi_{attn}(G_w(lex_C)) \label{eq:fergusr} \\
&\omega_{lex} = FC_1(concat(\omega_{C},~G_w(lex_p))) \nonumber \\
&\omega_{node} = RTN(\omega_{lex},~topology) \nonumber \\
&\Omega_{elementary} = FC_2(concat(repeat(\omega_{node}),~G_\Lambda(tagset_p))) \nonumber \\
&P_\theta(tag_{p,i}~|~RTN,~lex_p,~lex_{C}) = 
\frac{exp(\Psi_{predict}(\omega_{elementary_i})))}
{\sum_{j \in |\Omega|} exp(\Psi_{predict}(\omega_{elementary_j))}} \nonumber
\end{align}
%\vspace*{-0.45cm}

@@

\subsection{Linearization}
%\vspace{-0.1cm}

@@

\section{Experiments}
\label{sec:expt}
%\vspace*{-0.2cm}

@@

\subsection{Training}
%\vspace{-0.1cm}

@@

\subsection{Performance Metrics}
%\vspace{-0.1cm}

@@

\subsection{Results}
\label{sec:results}
%\vspace*{-0.1cm}

\begin{table}
\centering
\begin{tabular}{|l|p{3cm}|c|c|c|}
\cline{3-4}
\multicolumn{2}{}{}& \multicolumn{2}{ |c| }{ \textbf{Accuracy}}&\multicolumn{1}{}{} \\ \hline
\textbf{Model} & \textbf{Embedding}  & \textbf{Raw Model} 
& \textbf{After Decoding} & \textbf{Running Time} \\ \hline
Fergus-N &  Structural  &  58.17\% & 57.40\%  & 1.97s \\ \cline{2-5}
         &  Atomic      &  60.69\% & 55.56\% & 1.81s \\ \cline{2-5}
         &  Minimal     &  52.09\% & 54.18\% & 2.02s \\ \cline{2-5}
\hline
Fergus-R &  Structural &  67.62\% & 57.04\% & 0.30s \\ \cline{2-5}
         &  Atomic     &  82.65\% & 62.73\% & 0.36s\\ \cline{2-5}
         &  Minimal    & 10.13\% & 19.66\% & 0.54s \\ \cline{2-5}
\hline
\end{tabular}
\caption{For each supertag and embedding pair, the mean accuracy of
  supertag classification directly output by the model and in the
  consistent global assignment output by A* decoding. Also shown is
  the median running time---which includes model computation and A*
  search. The structural embeddings are computed with convolutional coding, the atomic embeddings as rows in a matrix, and the minimal embeddings as scalars in a vector.}
\label{table:accresults}
%\vspace*{-0.5cm}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{|l|p{3cm}|p{2.5cm}|r|}
\cline{3-4}
\multicolumn{2}{}{} & \multicolumn{2}{|c|}{Accuracy}   \\ \hline
\textbf{Model} & \textbf{Embedding}  & \textbf{Top Scoring} & \textbf{Best Performance} \\ \hline
Fergus-N & Structural & 65.80\% & 72.58\% \\ \cline{2-4}
         & Atomic       & 65.52\%  & 71.82\% \\ \cline{2-4}
         & Minimal & 63.79\% & 71.09\% \\ \cline{2-4}
\hline
Fergus-R & Structural & 68.22\% & 74.70\% \\ \cline{2-4}
         & Atomic       &  69.29\% & 75.56\% \\ \cline{2-4}
         & Minimal &  58.23\% & 65.04\% \\ \cline{2-4}
\hline
\end{tabular}
\caption{Shown above as accuracy is the percentage of tokens in the linearized strings that are in correct positions according to an edit distance measure.}
\label{table:linresults}
%\vspace*{-0.5cm}
\end{table}

@@

\section{Related Work}
\label{sec:relatedwork}
%\vspace*{-0.2cm}

@@

\section{Conclusion}
\label{sec:conclusion}
%\vspace*{-0.2cm}

@@

\section*{Acknowledgments}
%\vspace*{-0.2cm}

@@

\section{Appendix}
\label{sec:appendix}
%\vspace{-0.2cm}

\begin{table}[t!]
\centering
\begin{tabular}{lr}
\begin{tabular}[t]{|p{6cm}|r|}
\hline
\textbf{Model Parameter} & \textbf{Value} \\ \hline
\hline \multicolumn{2}{|c|}{\textbf{Fergus-N Parameters}} \\ \hline
Fully connected layer size & 256 \\ \hline
Batch size & 128 \\ \hline
\hline  \multicolumn{2}{|c|}{\textbf{Fergus-R Parameters}} \\ \hline
Fully connected layer size & 256 \\ \hline 
Hidden state size & 128 \\ \hline
Batch size & 16 \\ \hline
\hline \multicolumn{2}{|c|}{\textbf{Embedding Parameters}} \\ \hline
Convolution filter size & 48 \\ \hline
Syntactic category embedding size & 32 \\ \hline
Node role embedding size & 32 \\ \hline
Word embedding size \newline
\cite{pennington2014glove} & 300 \\ \hline
\end{tabular}
&
\begin{tabular}[t]{|p{6cm}|r|}
\hline
\textbf{Model Parameter} & \textbf{Value} \\ \hline
\hline \multicolumn{2}{|c|}{\textbf{Language Model Parameters}} \\ \hline
Hidden state size & 368 \\ \hline
Batch size & 32 \\ \hline
\hline \multicolumn{2}{|c|}{\textbf{Optimization Parameters}} \\ \hline
Optimization Algorithm & ADAM \\ \hline
Fergus-R and Fergus-N Learning Rate & 1e-4 \\ \hline
Language Model Learning Rate & 0.01 \\ \hline
Fully-Connected Dropout Rate & 0.5 \\ \hline
Recurrent Weight Dropout Rate & 0.2 \\ \hline
$L_2$ Weight Decay & 1e-6 \\ \hline
Max gradient norm & 10.0 \\ \hline
Gradient clip threshold & 5.0 \\ \hline
\end{tabular}
\end{tabular}
\caption{The parameters for the Fergus-R, Fergus-N, and language models.  The
exact specifications in configuration files can be found in the code repository that accompanies this paper.}
\label{tab:params}
%\vspace*{-0.5cm}
\end{table}

@@

\subsection{Data}
%\vspace*{-0.1cm}

@@

\subsection{Model parameters}
%\vspace*{-0.1cm}

@@

\subsection{Model design}
A subset of all supertags are used in supertag probability distribution calculations
%
They are the set of supertags that are seen paired with the lexical item.
%
This reduces the output distribution from over 5000 to under 500.  
%
During computations, indices to the supertag embedding matrix are passed into the computation with the rest of the data. 
%
While there are several ways the affinity between supertag and lexical state vector could have been computed, we chose to concatenate the vectors, map them to a new space using a fully connected layer, and compute a score with a vectorized function.
%
The vectorized function operation is the same mechanism which calculates the probability distribution used in soft attention.
